The Decreasing Orbital Period of WASP-4\lowercase{b}

### 2018-12-05 paper/ms.tex
cleanup TeX

\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}

@@

Long-term monitoring of hot Jupiter transit and occultation times
should eventually reveal two distinct processes.  First, whenever the
total spin and orbital angular momenta in a hot Jupiter system are
below a particular critical value, the system will be unstable to
tidal decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most
hot Jupiters satisfy this condition, so their orbits should be
shrinking \citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate
of decay is uncertain, and depends on how friction dissipates energy
carried by gravitational tides (\citealt{ogilvie_tidal_2014} gives a
review).  If we could directly measure orbital decay rates, it would
inform our understanding of tidal dissipation.  This is important
because the dissipation rates determine the fate of the system.
During its inspiral, if the planet fills its Roche lobe below the
stellar surface, the drag force from the stellar envelope accelerates
the inspiral, leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Otherwise, if the
planet fills its Roche lobe above above the stellar surface, it could
either be dynamically disrupted into an accretion disk, or it could
slowly transfer mass to the star \citep{metzger_optical_2012}.  More
broadly, the outcome of the eventual interaction between the Earth and
red giant Sun also hinges on the efficiency of tidal dissipation
\citep{rasio_tidal_1996}.

@@

% % The HJ semimajor axis distributions and rotation period distribn
% % indicate that tidal decay is real.  (So too, might the age
% % distribution.)
% Stars hosting hot Jupiters tend to spin faster than other stars of
% similar age and mass \citep{penev_hats-18_2016,penev_empirical_2018}.
% This is expected because when the orbital frequency is larger than the
% spin rotation frequency, tides donate angular momentum from the planet
% to the star.
% The reverse process occurs between the Earth and Moon, leading to the
% Moon's slowly increasing separation from Earth of $X.XX$cm~yr$^{-1}$
% (CITE lunar ranging experiment).
% 
% The hot Jupiter star-planet separation distribution is best reproduced
% by models that account for tides~\citep{cameron_hierarchical_2018}.

% There have been some indications of period timing changes for hot jupiters
The most convincing signs seen to date of either orbital decay or apsidal
precession have been in WASP-12b.
\citet{maciejewski_departure_2016} showed that the transit times of
WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
\citet{patra_2017} presented new transit and occultation times for
the system, agreed that curvature was present, and found
$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
% Both \citet{maciejewski_departure_2016} and \citet{patra_2017}
% emphasized that the timing variations could be caused by , or
% also by tidal decay.
Assuming that the period change was caused entirely by tidal decay,
\citet{patra_2017} found values for the stellar tidal quality factor
of $Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
explained by standard equilibrium tide
models~\citep{penev_tidal_2011,ogilvie_tidal_2014}.
\citet{weinberg_tidal_2017} found though that if WASP-12 were a
subgiant, tidal dissipation due to a dynamical tide might be
sufficient to produce the required level of friction.
\cite{bailey_understanding_2019} concurred, but noted that the
observational constraints on WASP-12 itself favor a main-sequence
star, not a subgiant.  Thus, more data are needed to determine the
cause of the timing variations in WASP-12b.

% % There are also indications that apsidal precession should happen. If
% % there are any trace remnants of eccentricity, this will produce
% % timing variations.
% Timing variations can also be caused by rotation of
% the orbital ellipse within the orbital plane (apsidal precession).
% This happens because any rotational or tidal bulge creates a
% gravitation quadrupole field ($r^{-3}$, rather than the typical
% $r^{-2}$), which causes the line of apsides to move forward or back.
% \citet{ragozzine_probing_2009} examined apsidal precession.

% Tidal decay of hot Jupiters has been invoked to explain certain
% properties of the ensemble of star-planet systems. For example, the
% scarcity of gas giants with periods less than a day is suggestive of
% orbital decay \citep[see,
%   e.g.][]{Jackson+2008,Hansen2010,Penev+2012,Ogilvie2014}. The
% anomalously rapid rotation of some hot-Jupiter host stars has been
% attributed to transfer of the planet's orbital angular momentum
% \citep{Penev+2016}. The absence of hot Jupiters around subgiant stars
% may be caused by an acceleration of orbital decay when a star leaves
% the main sequence
% \citep{VillaverLivio2009,Hansen2010,SchlaufmanWinn2013}.  Tidal decay
% might also be responsible for the lower occurrence of close-in planets
% around rapidly rotating stars \citep{TeitlerKonigl2014}, or the
% realignment of stars and their planetary orbits
% \citep{MatsakosKonigl2015}. However, direct evidence for orbital decay
% has been lacking: there have been no clear demonstrations of a
% long-term period decrease due to orbital decay \citep[see,
%   e.g.,][]{Hoyer+2016,Wilkins+2017}.
% 
% Another unfulfilled prediction is that the orbits of hot Jupiters
% should be apsidally precessing on a timescale of decades
% \citep{MiraldaEscude2002,HeylGladman2007,PalKocsis2008,JordanBakos2008},
% as long as the orbits are at least slightly eccentric. In particular,
% \citet{RagozzineWolf2009} noted that the theoretical precession rate
% is dominated by the contribution from the planet's tidally deformed
% mass distribution. They advocated a search for apsidal precession as
% a means of probing the interiors of hot Jupiters.
%

@@

Using the times, flux measurements, and errors determined above, we
began our measurement procedure by using the Box Least Squares (BLS)
method to estimate the transit epoch, period, and
duration~\citep{kovacs_box-fitting_2002}.
From the BLS parameters, we isolated each transit to a rough estimate
of its midtime, $\pm$10 transit durations.  In each transit window, we
then simultaneously fit for a model transit, plus a local linear
trend.  Our transit model uses the analytic formulae calculated by
\citet{mandel_analytic_2002}, and implemented by
\citet{kreidberg_batman_2015}.

@@

We fixed the remaining transit parameters as follows.  We set the
eccentricity to zero, and the longitude of periastron to $\pi/2$.  For
the limb darkening, we assumed a fixed quadratic law, with
coefficients interpolated from the tables computed
by~\citet{claret_limb_2017}.  For the period, $a/R_\star$, and
inclination, we adopted the values from~\citet{petrucci_no_2013}.
These are in reasonable agreement with the parameters reported by
\citet{gillon_improved_2009}, \citet{southworth_high-precision_2009},
and \citet{huitson_gemini_2017}.

Having defined all the free and fixed parameters, we calculated an
initial guess for the free parameters by maximizing the likelihood.
We then sampled over the posterior using the algorithm proposed
by~\citet{goodman_ensemble_2010}, and implemented
by~\citet{foreman-mackey_emcee_2013}.  During this process, we used
the \texttt{corner} software to check that our posteriors were being
adequately sampled~\citep{corner_2016}.

@@

Important contributions to the transit timing data are as follows.
The earliest epoch in our fit is provided by the EulerCam lightcurve
observed by \citet{wilson_wasp-4b_2008}.  The second, more precise,
epoch comes from $z$-band photometry acquired by
\citet{gillon_improved_2009} at the VLT with FORS2.
Subsequent epochs came from \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded the GPS clock
  time in UT reference. They then converted this to BJD in the TDB
  reference for their reported midtimes.},
\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011},
and \citet{nikolov_wasp-4b_2012}.
%\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS clock
%	time in UT reference. They then converted this to BJD in the TDB
%	reference for their reported midtimes.},
%and reported transit midtimes precise to
%6~seconds.  \citet{dragomir_terms_2011}
%updated the ephemeris using two transits measured with the CTIO 0.9-m
%and 1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
%transits from
%the Baade telescope at Las Campanas\footnote{
%\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
%to show that the star's rotation axis is nearly aligned with the
%planet's orbital axis, in agreement with the Rossiter-McLaughlin
%measurement by \citet{triaud_spin-orbit_2010}.}.
%\citet{nikolov_wasp-4b_2012} observed 3 transits, in 4 independent
%photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
%the MPG/EOS-2.2~m at La Silla.
\citet{hoyer_tramos_2013} reported
nine transit observations taken with the the Y4KCAM on the SMARTS 1-m
telescope, and another three with the SOI on the SOAR 4.2-m telescope.
\citet{hoyer_tramos_2013} also homogeneously
redetermined the transit times from all previous studies.  Though
their latest two epochs did fall slightly below the prediction of a
linear ephemeris (Panel C of their Figure~6), they did not find any
statistical need for a quadratic function to fit the O-C values.

@@

There have been other studies of WASP-4b for which the transit timing
data are either hetereogenous, not available, or not useable.  We
omitted the timing measurements from
\citet{southworth_high-precision_2009}, since there were
technical problems with the computer clock at the time of
observation~\citep{nikolov_wasp-4b_2012}.  The timing study
by~\citet{petrucci_no_2013} ruled out transit timing variations larger
than 54~seconds over the previous 5 years, but did not report transit
times.  \citet{baluev_benchmarking_2015} collected transit times for
many hot Jupiters, including WASP-4b, and included times from the
Exoplanet Transit Database
(ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
\citep{poddany_ETD_2010}.  
Since the ETD data come from heterogeneous sources, their timestamps
are less clearly documented, and the times are thus more prone to
systematic errors, we omit them from consideration.
%We consider the effects of including ETD data in Appendix~\ref{sec:verify_archival_times}.
\citet{may_mopss_2018} obtained optical spectra using IMCAS on
Magellan, and found a flat transmission spectrum, in agreement with
the results reported by \citet{huitson_gemini_2017}.  They did not
report transit times.

@@

Figure~\ref{fig:times} shows the residuals with respect to the linear
model.
The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of freedom.
The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57 degrees of freedom.
The median MCMC precession fit has $\chi^2 = 51.0$ and 55 degrees of freedom.

%FIXME UPDATE THESE NUMBERS!!

@@

The quadratic model fits better than the precession model.
It is favored over the precession fit by $\Delta \chi^2 = 3.0$,
and has two fewer free parameters.
Another heuristic useful for model comparison is the Bayesian
Information Criterion (BIC),
\begin{equation}
  {\rm BIC} = \chi^2 + k\log n,
\end{equation}
for $k$ the number of free parameters, and $n$ the number of data
points. For us, $n=60$.
The difference in the BIC for the precession and decay models is
$\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm BIC}_{\rm quad} = 11$,
corrsonding to a Bayes factor of $\approx 4\times10^{4}$.
In the language of~\citet{kass_bayes_1995}, this is ``decisive
evidence'' for the quadratic model describing the data better than the
precession model.
Using the Akaike Information Criterion similarly favors the orbital
decay model, by $\Delta {\rm AIC} = 7$.

Assuming the orbital decay model, we find a period derivative 
\begin{equation}
\dot{P}
    = - (3.82^{+0.38}_{-0.37})\times 10^{-10}
    = - (12.1^{+1.2}_{-1.2})\,{\rm ms}\,{\rm yr}^{-1}.
\end{equation}
For comparison, the period derivative seen by
\citet{maciejewski_departure_2016} and \citet{patra_2017} in
WASP-12b is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
Assuming the quadratic model, WASP-4b is thus decaying at
about half the rate as WASP-12b.

If we instead assume the precession model, the best-fit
eccentricity is $e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}$. The
longitude of periapse correspondingly advances by $\dot{\omega}
= 14.8^{+5.4}_{3.8}\ {\rm degrees}\,{\rm yr}^{-1}$.   The
corresponding precession period would be $24^{+8}_{-6}$ years.

@@

Let us assume further that the 
``constant phase lag'' model for tidal interaction in
a binary applies \citep{zahn_tidal_1977}.
%, the travelling equilibrium tidal
%bulge is damped by turbulent viscosity in the convective envelope.
The fluid response is parametrized by a constant modified\footnote{For
stars, $k_\star \sim \mathcal{O}(10^{-2})$, so it is important to
explicitly distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
e.g.},][]{schwarzschild_structure_1958}.} quality factor, $Q_\star' =
3 Q_\star / (2k_\star)$.  Here $k_\star$ is the stellar Love number,
which is smaller when the star's density distribution is more
centrally concentrated. $Q_\star$ is the ratio of the
energy stored in the equilibrium deformation of the star, divided by
the energy dissipated per tidal period \citep[{\it
e.g.},][]{goldreich_q_1966}.  A larger $Q_\star$ implies less
efficient dissipiation.  Though the fluid mechanics responsible for
the dissipation should depend on the amplitude and frequency of
the tidal perturbation \citep[][Section~3.3]{ogilvie_tidal_2014}, we
will treat $Q_\star'$ as a constant.  This should not be taken
literally; $Q_\star'$ may in fact change sharply with varying tidal
frequency \citep{penev_empirical_2018}.

@@

Occultation timing can provide stronger eccentricity limits than
RV measurements, but they often come with caveats.
For WASP-4b, the strongest limit on its eccentricity comes from the
secondary eclipses observed by \citet{beerer_secondary_2011}, which
led to a $2\sigma$ upper limit on $| e\cos\omega | $ of 0.0024.
% \citet{caceres_ground-based_2011} report $e\cos\omega = 0.0027 \pm
% 0.0018$. 
% \citet{zhou_secondary_2015} report a weaker $e\cos\omega = -0.001
% \pm 0.003$. 
Placing eccentricity limits stronger than $e< 10^{-3}$
has only been possible for a few hot Jupiters, and often with numerous
false-starts.  HD 189733b, for example, was initially reported to have
a very small eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
\citep{knutson_map_2007}.  This result was retracted by
\citet{agol_climate_2010}, who used a larger dataset of 7 Spitzer
eclipses and found a smaller average offset in the secondary eclipse
time than \citet{knutson_map_2007}.  \citet{agol_climate_2010} also
noted that a larger fraction of the planet's observed occultation time
offset could be explained by detailed modeling of the planetary
hot-spot, which can change occultation time offsets by up to a few
minutes \citep{williams_resolving_2006}.  Including an updated
hot-spot model, \citet{agol_climate_2010} reported a 2$\sigma$ limit
of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit
yet for a hot Jupiter.  A similar saga occurred for HD 209458b, for which
\citet{winn_measurement_2005} combined Spitzer secondary eclipses with
RV data and reported a best-fit eccentricity of $e>0.0057$ at 90\%
confidence.  They suspected that the data were consistent with zero
eccentricity, and that the lower bound on $e$ was an artifact of
imperfect limb-darkening model.  Stronger limits were eventually put
in place by \citet{crossfield_spitzer_mips_2012}, who found
$e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b,
for which the discovery RVs and photometry led to a non-zero but small
eccentricity measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.
A prompt ground-based occultation measurement by
\citet{lopez-morales_day-side_2010} concurred with weak statistical
significance that a small eccentricity was favored, but
further follow-up found occultation times consistent with circular
orbits; $| e\cos\omega| < 0.0040 $
\citep{campo_orbit_2011,croll_near-infrared_2011}.

@@

Another mechanism to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the companion would need to be
inclined relative to that of the hot Jupiter by at least $\sin^{-1}
\sqrt{2/5} \approx 39^\circ$.
\citet{bailey_understanding_2019} point out how the joint requirements
of Kozai-Lidov oscillations with a non-detection of an additional RV
signal can be combined to produce lower and upper limits on the outer
companion's mass and the semimajor axis ratio, $a_{\rm c}/a_{\rm b}$.
For WASP-4, ...
%TODO: check this mechanism
%FIXME: take a look at RVs. rule this shit out. /  understand where
%the bounds are.

>>>>

\correspondingauthor{L. G. Bouma}
\email{luke@astro.princeton.edu}

@@

Long-term monitoring of hot Jupiter transit and occultation times
should eventually reveal two distinct processes.  First, whenever the
summed spin and orbital angular momenta in a hot Jupiter system are
below a particular critical value, the system will be unstable to
tidal decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most
hot Jupiters satisfy this condition, so their orbits should be
shrinking \citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate
of decay is uncertain, and depends on how friction dissipates energy
carried by gravitational tides (\citealt{ogilvie_tidal_2014} gives a
review).  If we could directly measure orbital decay rates, it would
inform our understanding of tidal dissipation.  This is important
because the dissipation rates determine the fate of the system.
During its inspiral, if the planet fills its Roche lobe below the
stellar surface, the drag force from the stellar envelope accelerates
the inspiral, leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
the planet fills its Roche lobe above above the stellar surface, then
it could either be dynamically disrupted into an accretion disk or it
could slowly transfer mass to the star \citep{metzger_optical_2012}.
More broadly, the outcome of the eventual interaction between the
Earth and red giant Sun also hinges on the efficiency of tidal
dissipation \citep{rasio_tidal_1996}.

@@

% Stars hosting hot Jupiters tend to spin faster than other stars of
% similar age and mass
% \citep{penev_hats-18_2016,penev_empirical_2018}.  This is expected
% because when the orbital frequency is larger than the spin rotation
% frequency, tides donate angular momentum from the planet to the
% star.  The reverse process occurs between the Earth and Moon,
% leading to the Moon's slowly increasing separation from Earth of
% $X.XX$cm~yr$^{-1}$ (CITE lunar ranging experiment).  The hot Jupiter
% star-planet separation distribution is best reproduced by models
% that account for tides~\citep{cameron_hierarchical_2018}.

The most convincing signs seen to date of either orbital decay or apsidal
precession have been in WASP-12b.
\citet{maciejewski_departure_2016} showed that the transit times of
WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
\citet{patra_2017} presented new transit and occultation times for
the system, agreed that curvature was present, and found
$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
Assuming that the period change was caused entirely by tidal decay,
\citet{patra_2017} found values for the stellar tidal quality factor
of $Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
explained by standard equilibrium tide
models~\citep{penev_tidal_2011,ogilvie_tidal_2014}.
\citet{weinberg_tidal_2017} found though that if WASP-12 were a
subgiant, tidal dissipation due to a dynamical tide might be
sufficient to produce the required level of friction.
\cite{bailey_understanding_2019} concurred, but noted that the
observational constraints on WASP-12 itself favor a main-sequence
star, not a subgiant.  Thus, more data are needed to determine the
cause of the timing variations in WASP-12b.

% Tidal decay of hot Jupiters has been invoked to explain certain
% properties of the ensemble of star-planet systems. For example, the
% scarcity of gas giants with periods less than a day is suggestive of
% orbital decay \citep[see,
%   e.g.][]{Jackson+2008,Hansen2010,Penev+2012,Ogilvie2014}. The
% anomalously rapid rotation of some hot-Jupiter host stars has been
% attributed to transfer of the planet's orbital angular momentum
% \citep{Penev+2016}. The absence of hot Jupiters around subgiant stars
% may be caused by an acceleration of orbital decay when a star leaves
% the main sequence
% \citep{VillaverLivio2009,Hansen2010,SchlaufmanWinn2013}.  Tidal decay
% might also be responsible for the lower occurrence of close-in planets
% around rapidly rotating stars \citep{TeitlerKonigl2014}, or the
% realignment of stars and their planetary orbits
% \citep{MatsakosKonigl2015}. However, direct evidence for orbital decay
% has been lacking: there have been no clear demonstrations of a
% long-term period decrease due to orbital decay \citep[see,
%   e.g.,][]{Hoyer+2016,Wilkins+2017}.
%

@@

Using the times, flux measurements, and errors determined above, we
began our measurement procedure by using the Box Least Squares (BLS)
method to estimate the transit epoch, period, and duration
\citep{kovacs_box-fitting_2002}.  From the BLS parameters, we isolated
each transit to a rough estimate of its midtime, $\pm$10 transit
durations.  In each transit window, we then simultaneously fit for a
model transit, plus a local linear trend.  Our transit model uses the
analytic formulae calculated by \citet{mandel_analytic_2002}, and
implemented by \citet{kreidberg_batman_2015}.

@@

We fixed the remaining transit parameters as follows.  We set the
eccentricity to zero, and the longitude of periastron to $\pi/2$.  For
the limb darkening, we assumed a fixed quadratic law, with
coefficients interpolated from the tables computed by
\citet{claret_limb_2017}.  For the period, $a/R_\star$, and
inclination, we adopted the values from~\citet{petrucci_no_2013}.
These are in reasonable agreement with the parameters reported by
\citet{gillon_improved_2009}, \citet{southworth_high-precision_2009},
and \citet{huitson_gemini_2017}.

Having defined all the free and fixed parameters, we calculated an
initial guess for the free parameters by maximizing the likelihood.
We then sampled over the posterior using the algorithm proposed by
\citet{goodman_ensemble_2010}, and implemented by
\citet{foreman-mackey_emcee_2013}.  During this process, we used the
\texttt{corner} software to check that our posteriors were being
adequately sampled~\citep{corner_2016}.

@@

Important contributions to the transit timing data are as follows.
The earliest epoch in our fit is provided by the EulerCam lightcurve
observed by \citet{wilson_wasp-4b_2008}.  The second, more precise,
epoch comes from $z$-band photometry acquired by
\citet{gillon_improved_2009} at the VLT with FORS2.  Subsequent epochs
came from \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009}
recorded the GPS clock time in UT reference. They then converted this
to BJD in the TDB reference for their reported midtimes.},
\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
\citet{nikolov_wasp-4b_2012}.
%\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS
%clock time in UT reference. They then converted this to BJD in the
%TDB reference for their reported midtimes.}, and reported transit
%midtimes precise to 6~seconds.  \citet{dragomir_terms_2011} updated
%the ephemeris using two transits measured with the CTIO 0.9-m and
%1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
%transits from the Baade telescope at Las Campanas\footnote{
%\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
%to show that the star's rotation axis is nearly aligned with the
%planet's orbital axis, in agreement with the Rossiter-McLaughlin
%measurement by \citet{triaud_spin-orbit_2010}.}.
%\citet{nikolov_wasp-4b_2012} observed 3 transits, in 4 independent
%photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
%the MPG/EOS-2.2~m at La Silla.
\citet{hoyer_tramos_2013} reported nine transit observations taken
with the the Y4KCAM on the SMARTS 1-m telescope, and another three
with the SOI on the SOAR 4.2-m telescope.  \citet{hoyer_tramos_2013}
also homogeneously redetermined the transit times from all previous
studies.  Though their latest two epochs did fall slightly below the
prediction of a linear ephemeris (Panel C of their Figure~6), they did
not find any statistical need for a quadratic function to fit the O-C
values.

@@

There have been other studies of WASP-4b for which the transit timing
data are either hetereogenous, not available, or not useable.  We
omitted the timing measurements from
\citet{southworth_high-precision_2009}, since there were technical
problems with the computer clock at the time of
observation~\citep{nikolov_wasp-4b_2012}.  The timing study
by~\citet{petrucci_no_2013} ruled out transit timing variations larger
than 54~seconds over the previous 5 years, but did not report transit
times.  \citet{baluev_benchmarking_2015} collected transit times for
many hot Jupiters, including WASP-4b, and included times from the
Exoplanet Transit Database
(ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
\citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
sources, their timestamps are less clearly documented, and the times
are thus more prone to systematic errors, we omit them from
consideration.
%We consider the effects of including ETD data in
%Appendix~\ref{sec:verify_archival_times}.
\citet{may_mopss_2018} obtained optical spectra using IMCAS on
Magellan, and found a flat transmission spectrum, in agreement with
the results reported by \citet{huitson_gemini_2017}.  They did not
report transit times.

@@

Figure~\ref{fig:times} shows the residuals with respect to the linear
model.
%TODO: final numbers
The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of freedom.
The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57 degrees of freedom.
The median MCMC precession fit has $\chi^2 = 51.0$ and 55 degrees of freedom.

@@

%TODO: final numbers
The quadratic model fits better than the precession model.
It is favored over the precession fit by $\Delta \chi^2 = 3.0$,
and has two fewer free parameters.
Another heuristic useful for model comparison is the Bayesian
Information Criterion (BIC),
\begin{equation}
  {\rm BIC} = \chi^2 + k\log n,
\end{equation}
for $k$ the number of free parameters, and $n$ the number of data
points. For us, $n=60$.
The difference in the BIC for the precession and decay models is
$\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm BIC}_{\rm quad} = 11$,
corrsonding to a Bayes factor of $\approx 4\times10^{4}$.
In the language of~\citet{kass_bayes_1995}, this is ``decisive
evidence'' for the quadratic model describing the data better than the
precession model.
Using the Akaike Information Criterion similarly favors the orbital
decay model, by $\Delta {\rm AIC} = 7$.

%TODO: final numbers
Assuming the orbital decay model, we find a period derivative 
\begin{equation}
\dot{P}
    = - (3.82^{+0.38}_{-0.37})\times 10^{-10}
    = - (12.1^{+1.2}_{-1.2})\,{\rm ms}\,{\rm yr}^{-1}.
\end{equation}
For comparison, the period derivative seen by
\citet{maciejewski_departure_2016} and \citet{patra_2017} in
WASP-12b is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
Assuming the quadratic model, WASP-4b is thus decaying at
about half the rate as WASP-12b.

%TODO: final numbers
If we instead assume the precession model, the best-fit
eccentricity is $e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}$. The
longitude of periapse correspondingly advances by $\dot{\omega}
= 14.8^{+5.4}_{3.8}\ {\rm degrees}\,{\rm yr}^{-1}$.   The
corresponding precession period would be $24^{+8}_{-6}$ years.

@@

Let us assume further that the ``constant phase lag'' model for tidal
interaction in a binary applies \citep{zahn_tidal_1977}.  The fluid
response is parametrized by a constant modified\footnote{For stars,
$k_\star \sim \mathcal{O}(10^{-2})$, so it is important to explicitly
distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
e.g.},][]{schwarzschild_structure_1958}.} quality factor, $Q_\star' =
3 Q_\star / (2k_\star)$.  Here $k_\star$ is the stellar Love number,
which is smaller when the star's density distribution is more
centrally concentrated. $Q_\star$ is the ratio of the energy stored in
the equilibrium deformation of the star, divided by the energy
dissipated per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.
A larger $Q_\star$ implies less efficient dissipiation.  Though the
fluid mechanics responsible for the dissipation should depend on the
amplitude and frequency of the tidal perturbation
\citep[][Section~3.3]{ogilvie_tidal_2014}, we will treat $Q_\star'$ as
a constant.  This should not be taken literally; $Q_\star'$ may in
fact change sharply with varying tidal frequency
\citep{penev_empirical_2018}.

@@

Occultation timing can provide stronger eccentricity limits than RV
measurements, but they often come with caveats.  For WASP-4b, the
strongest limit on its eccentricity comes from the secondary eclipses
observed by \citet{beerer_secondary_2011}, which led to a $2\sigma$
upper limit on $| e\cos\omega | $ of 0.0024.
% \citet{caceres_ground-based_2011} report $e\cos\omega = 0.0027 \pm
% 0.0018$.  \citet{zhou_secondary_2015} report a weaker $e\cos\omega =
% -0.001 \pm 0.003$. 
Placing eccentricity limits stronger than $e< 10^{-3}$ has only been
possible for a few hot Jupiters, and often with numerous false-starts.
HD 189733b, for example, was initially reported to have a very small
eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
\citep{knutson_map_2007}.  This result was retracted by
\citet{agol_climate_2010}, who used a larger dataset of 7 Spitzer
eclipses and found a smaller average offset in the secondary eclipse
time than \citet{knutson_map_2007}.  \citet{agol_climate_2010} also
noted that a larger fraction of the planet's observed occultation time
offset could be explained by detailed modeling of the planetary
hot-spot, which can change occultation time offsets by up to a few
minutes \citep{williams_resolving_2006}.  Including an updated
hot-spot model, \citet{agol_climate_2010} reported a 2$\sigma$ limit
of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit
yet for a hot Jupiter.  A similar saga occurred for HD 209458b, for
which \citet{winn_measurement_2005} combined Spitzer secondary
eclipses with RV data and reported a best-fit eccentricity of
$e>0.0057$ at 90\% confidence.  They suspected that the data were
consistent with zero eccentricity, and that the lower bound on $e$ was
an artifact of imperfect limb-darkening model.  Stronger limits were
eventually put in place by \citet{crossfield_spitzer_mips_2012}, who
found $e\cos\omega < 0.0007$ at $2\sigma$.  The final example is
WASP-12b, for which the discovery RVs and photometry led to a non-zero
but small eccentricity measurement \citep[][at
$3\sigma$]{hebb_wasp-12b_2009}.  A prompt ground-based occultation
measurement by \citet{lopez-morales_day-side_2010} concurred with weak
statistical significance that a small eccentricity was favored, but
further follow-up found occultation times consistent with circular
orbits; $| e\cos\omega| < 0.0040 $
\citep{campo_orbit_2011,croll_near-infrared_2011}.

@@

Another mechanism to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the companion would need to be
inclined relative to that of the hot Jupiter by at least $\sin^{-1}
\sqrt{2/5} \approx 39^\circ$.
\citet{bailey_understanding_2019} point out how the joint requirements
of Kozai-Lidov oscillations with a non-detection of an additional RV
signal can be combined to produce lower and upper limits on the outer
companion's mass and the semimajor axis ratio, $a_{\rm c}/a_{\rm b}$.
For WASP-4, ...
% TODO: check this mechanism
% FIXME: take a look at RVs. rule this out, or understand where the
% bounds are.
